A Parallelized and Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo

### 2021-07-19 latex_files/ch2/2_4_ROMC.tex
updated most content in chapter 2,3
initial draft (without intro and conclusion)

The simplifications introduced by OMC, although quite useful from a
computational point-of-view, they suffer from some significant failure modes:

\begin{itemize}
\item The whole acceptable region $\mathcal{S}_i$, for each nuisance variable,
  shrinks to a single point $\thetab_i^*$; this simplification may add
  significant error when then the area $\mathcal{S}_i$ is relatively big.
\item The weight $w_i$ is computed based only at the curvature at the
  point $\thetab_i^*$. This approach is error prone at many cases
  e.g.\ when $g_i$ is almost flat at $\thetab_i^*$, leading to a
  $\text{det}(\jac_i^{*T}\jac_i^*) \rightarrow 0 \Rightarrow w_i
  \rightarrow \infty$, thus dominating the posterior.
\item There is no way to solve the optimisation problem
  $\thetab_i^* = \text{argmin}_{\thetab} \: g_i(\thetab)$ when $g_i$
  is not differentiable.
\end{itemize}

@@

The ROMC approach resolves the aforementioned issues. Instead of
collapsing the acceptance regions $\mathcal{S}_i$ into single points,
it tries to approximate them with a bounding box.\footnote{The
  description on how to estimate the bounding box is provided in the
  following chapters.}. A uniform distribution is then defined on the
bounding box area, used as the proposal distribution for importance
sampling. If we define as $q_i$, the uniform distribution defined on
the $i-th$ bounding box, weighted sampling is performed as:

\begin{gather}
  \label{eq:sampling}
  \thetab_{ij} \sim q_i \\
  w_{ij} = \frac{\indicator{\region(\data)}(M_d(\thetab_{ij}, \vb_i)) p(\thetab_{ij})}{q(\thetab_{ij})}
\end{gather}

\noindent
Having defined the procedure for obtaining weighted samples, any
expectation $E_{p(\thetab|\data)}[h(\thetab)]$, can be approximated
as,

@@

\subsubsection{Construction of the proposal region}

In this section we will describe mathematically the steps needed for
computing the proposal distributions $q_i$. There will be also
presented a Bayesian optimisation alternative when gradients are not
available.

\subsubsection*{Define and solve deterministic optimisation problems}

For each set of nuisance variables $\vb_i, i = \{1,2,\ldots,n_1 \}$ a
deterministic function is defined as
$f_i(\thetab) = M_d(\thetab,\vb_i)$. For constructing the proposal
region, we search for a point
$\thetab_* : d(f_i(\thetab_*), \data) < \epsilon$; this point can be
obtained by solving the the following optimisation problem:

\begin{subequations}
\begin{alignat}{2}
&\!\min_{\thetab}        &\qquad& g_i(\thetab) = d(\data,  f_i(\thetab))\label{eq:optProb}\\
&\text{subject to} &      & g_i(\thetab) \leq \epsilon
\end{alignat}
\end{subequations}
%
We maintain a list of the solutions $\thetab_i^*$ of the optimisation
problems. If for a specific set of nuisance variables $\vb_i$, there
is no feasible solution we add nothing to the list. The optimisation
problem can be treated as unconstrained, accepting the optimal point
$\thetab_i^* = \text{argmin}_{\thetab} g_i(\thetab)$ only if
$g_i(\thetab_i^*) < \epsilon$.

\subsubsection*{Gradient-based approach}
\label{subsubsec:GB_approach}

The nature of the generative model $M_r(\thetab)$, specifies the
properties of the objective function $g_i$. If $g_i$ is continuous
with smooth gradients $\nabla_{\thetab} g_i$ any gradient-based
iterative algorithm can be used for solving~\ref{eq:optProb}. The
gradients $\nabla_{\thetab} g_i$ can be either provided in closed form
or approximated by finite differences.

\subsubsection*{Bayesian optimisation approach}
\label{subsubsec:GP_approach}

In cases where the gradients are not available, the Bayesian
optimisation scheme provides an alternative
choice~\cite{Shahriari2016}. With this approach, apart from
obtaining an optimal $\thetab_i^* $, a surrogate model $\hat{d}_i$ of
the distance $g_i$ is fitted; this approximate model can be used in
the following steps for making the method more
efficient. Specifically, in the construction of the proposal region
and in
equations~\eqref{eq:approx_posterior},~\eqref{eq:sampling},~\eqref{eq:expectation}
it could replace $g_i$ in the evaluation of the indicator
function, providing a major speed-up.

\subsubsection*{Construction of the proposal area $q_i$}

After obtaining a $\thetab_i^*$ such that
$g_i(\thetab_i^*) < \epsilon$, we need to construct a bounding box
around it. The bounding box $\mathcal{\hat{S}}_i$ must contain the
acceptance region around $\thetab_i^*$, i.e.\
$\{ \thetab : g_i(\thetab) < \epsilon$,
$d(\thetab, \thetab_i^*) < M \}$. The second condition
$d(\thetab, \thetab_i^*) < M$ is meant to describe that if
$\mathcal{S}_i := \{ \thetab : g_i(\thetab) < \epsilon \} $ contains a
number of disjoint sets of $\thetab$ that respect
$g_i(\thetab) < \epsilon$, we want our bounding box to fit only the
one that contains $\thetab_i^*$. We seek for a bounding box that is as
tight as possible to the local acceptance region (enlarging the
bounding box without a reason decreases the acceptance rate) but large
enough for not discarding accepted areas.

In contrast with the OMC approach, we construct the bounding box by
obtaining search directions and querying the indicator function as we move on them. The
search directions $\mathbf{v}_d$ are computed as the eigenvectors of
the curvature at $\thetab_i^*$ and a line-search method is used to
obtain the limit point where
$g_i(\thetab_i^* + \kappa \vb_d) \geq
\epsilon$\footnote{$-\kappa$ is used as well for the opposite direction along the search line}. The Algorithm~\ref{alg:region_construction} describes the method in-depth. After the limits are obtained along all
search directions, we define bounding box and the uniform distribution $q_i$. This is the proposal distribution used for the importance
sampling as explained in \eqref{eq:sampling}.

@@

After the construction of the bounding box $\mathcal{\hat{S}}_i$, we
are no longer interested in the surface outside the box. In the future
steps (e.g. sampling, evaluating the posterior) we will only evaluate
$g_i$ inside the corresponding bounding box. Hence, we could fit a
local surrogate model $\hat{g}_i$ for representing the local area
around $\theta_i^*$. Doing so, in the future steps we can exploit
$\hat{g}_i$ for evaluating the indicator function instead of running
the whole deterministic simulator.

Any ML regression model may be chosen as local surrogates. The choice
should consider the properties of the local region (i.e. size,
smoothness). The ROMC proposes fitting a simple quadratic model. The
training set $X: \mathbb{R}^{N \times D}$ is created by sampling $N$
points from the bounding box and the labels $Y: \mathbb{R}^{N}$ are
the computed by evaluating $g_i$. The quadratic model is fitted on the
data points, for minimising the square error.

This additional step places an
additional step in the training part, increasing the computational
demands, but promises a major speed at the inference phase (sampling,
posterior evaluation). It is frequent in ML problems, to be quite
generous with the execution time at the training phase, but quite
eager at the inference phase. Fitting a local surrogate model aligns
with this requirement.

>>>>

The simplifications introduced by OMC, although beneficial from a
computational point-of-view, they suffer from some drawbacks:

\begin{itemize}
\item The whole acceptable region $\mathcal{S}_i$
  shrinks to a single point $\thetab_i^*$/. This simplification adds
  significant apprimation error when then the area $\mathcal{S}_i$ is relatively big.
\item The weight $w_i$ is computed based only at the curvature at the
  point $\thetab_i^*$. Using only local information for estimating the weight can also be misleading. For example, if $g_i$ is almost flat at $\thetab_i^*$, i.e. $\text{det}(\jac_i^{*T}\jac_i^*) \rightarrow 0, then \Rightarrow w_i
  \rightarrow \infty$ dominating the posterior.
\item There is no way to solve the optimisation problem
  $\thetab_i^* = \text{argmin}_{\thetab} \: g_i(\thetab)$ when $g_i$
  is not differentiable.
\end{itemize}

@@

ROMC resolves the failure-modes of OMC mainly by approaching the
acceptance regions $\mathcal{S}_i$ in a more sophisticated
way. Instead of collapsing the acceptance regions into single points,
we approximate them with a bounding box region. A uniform distribution
defined on the bounding box area, is then used as the \emph{proposal
  distribution} for importance sampling. If $q_i$ is the uniform
distribution defined on the $i-th$ bounding box, weighted sampling is
performed by drawing samples from $q_i$ and assigning the weights as
in~\eqref{eq:sampling}.

\begin{gather}
  \thetab_{ij} \sim q_i \\  \label{eq:sampling}
  w_{ij} = \frac{\indicator{\region(\data)}(M_d(\thetab_{ij}, \vb_i)) p(\thetab_{ij})}{q(\thetab_{ij})}
\end{gather}

\noindent

Having defined the procedure for obtaining weighted samples, any
expectation $E_{p(\thetab|\data)}[h(\thetab)]$, can be approximated
as,

@@

\subsubsection{Proposal region estimation}

The constrution of the proposal regions $q_i$ is performed in two
steps; (a) solving the deterministic optimisation problems as in OMC
and (b) constructing the bounding box regions.

\subsubsection*{Solving the deterministic optimisation problems}

For each set of nuisance variables $\vb_i, i = \{1,2,\ldots,n_1 \}$ a
deterministic function is defined as
$f_i(\thetab) = M_d(\thetab,\vb_i)$. We look for a point
$\thetab_* : d(f_i(\thetab_*), \data) < \epsilon$. This point can be
obtained by solving the the following optimisation problem:

\begin{subequations}
\begin{alignat}{2}
&\!\min_{\thetab}        &\qquad& g_i(\thetab) = d(\data,  f_i(\thetab))\label{eq:optProb}\\
&\text{subject to} &      & g_i(\thetab) \leq \epsilon
\end{alignat}
\end{subequations}
%
The optimisation problem can be treated as unconstrained, accepting
the optimal point $\thetab_i^* = \text{argmin}_{\thetab} g_i(\thetab)$
only if $g_i(\thetab_i^*) < \epsilon$. The nature of the generative
model $M_r(\thetab)$, specifies the properties of the objective
function $g_i$. If $g_i$ is differentiable any gradient-based
iterative algorithm can be used for~\ref{eq:optProb}. The gradients
$\nabla_{\thetab} g_i$ can be either provided in closed form or
approximated by finite differences. Alternatively, if $g_i$ is not
differentiable, the Bayesian Optimisation approach provides an
alternative choice~\cite{Shahriari2016}. In this case, apart from
obtaining an optimal $\thetab_i^* $, a surrogate model $\hat{d}_i$ of
the distance $g_i$ is also fitted; the approximate model can then
substitute the actual model in the following steps, e.g. construction
of bounding box, sampling, for accelerating the method.

\subsubsection*{Bounding boxes construction}

After obtaining a $\thetab_i^*$ such that
$g_i(\thetab_i^*) < \epsilon$, we need to construct a bounding box
area $\mathcal{\hat{S}}_i$ around it. The bounding box must contain
the acceptance region around $\thetab_i^*$, formally
$\{ \thetab : g_i(\thetab) < \epsilon$,
$d(\thetab, \thetab_i^*) < M \}$. The second condition
$d(\thetab, \thetab_i^*) < M$ describes that if
$\mathcal{S}_i := \{ \thetab : g_i(\thetab) < \epsilon \} $ contains a
number of disjoint sets of $\thetab$ that they fulfill
$g_i(\thetab) < \epsilon$, we want the bounding box to match only the
set that contains $\thetab_i^*$. We seek for a bounding box that is as
tight as possible to the local acceptance region for ensuring high
acceptance rate in the importance sampling.

Bounding boxes are built in two steps. Firstly, the search directions
$\mathbf{v}_d$ are computed as the eigenvectors of the curvature at
$\thetab_i^*$ and then the limits are obtained with a line-search
method looking for the smallest
$\kappa: g_i(\thetab_i^* + \kappa \vb_d) \geq
\epsilon$\footnote{$-\kappa$ is used as well for the opposite
  direction along the search line}. The
Algorithm~\ref{alg:region_construction} describes the method
in-depth. After limits are found along all directions the uniform
distribution $q_i$ can be defined and is used as the proposal region
for importance sampling.

@@

After the construction of the bounding box $\mathcal{\hat{S}}_i$, we
are no longer interested in $g_i(\thetab)$ outside
$\mathcal{\hat{S}}_i$. In the inference steps (e.g. sampling,
evaluating the posterior), we evaluate $g_i$ for checking if a sample
from the bounding box belongs to the acceptance region. Instead of
running the deterministic simulator $M_d(\thetab_{ij}, \vb_i)$ for
evaluating $g_i$, ROMC proposes fitting a \emph{light} surrogate model
$\hat{g}_i$ for approximating $g_i$ around $\theta_i^*$.

ROMC proposes fitting a simple quadratic model\footnote{Any ML
  regression model may be chosen as local surrogate. The choice should
  consider the properties of the local region (i.e. size,
  smoothness). } by sampling $N$ points from the bounding box and the
labels $Y: \mathbb{R}^{N}$ are computed by evaluating $g_i$. This
additional step adds an extra effort in the training part, but
promises a significant speed-up at the inference phase. Such imbalance
is usual in the context of ML where is usual to be quite generous with
the execution time at the training phase but quite eager at the
inference phase. Fitting a local surrogate model aligns with this
requirement.

### 2021-07-19 latex_files/ch2/2_5_ROMC_algorithmic.tex
initial draft (without intro and conclusion)

In this section, we will provide the algorithmic description of the
ROMC method; how to solve the optimisation problems using either the
gradient-based approach or the Bayesian optimisation alternative and
the construction of the bounding box. Afterwards, we will discuss the
advantages and disadvantages of each choice in terms of accuracy and
efficiency.

At a high-level, the ROMC method can be split into the training and
the inference part.

\subsubsection*{Training part}
\noindent
At the training (fitting) part, the goal is the estimation of the
proposal regions $q_i$. The tasks are (a) sampling the nuisance
variables $\vb_i \sim p(\vb)$ (b) defining the optimisation problems
$\min_{\thetab} \: g_i(\thetab)$ (c) obtaining $\thetab_i^*$ (d)
checking whether $d_i^* \leq \epsilon$ and (e) building the bounding
box for obtaining the proposal region $q_i$. If gradients are
available, using a gradient-based method is advised for obtaining
$\thetab_i^*$ much faster. Providing $\nabla_{\thetab} g_i$ in
closed-form provides an upgrade in both accuracy and efficiency; If
closed-form description is not available, approximate gradients with
finite-differences requires two evaluations of $g_i$ for
\textbf{every} parameter $\thetab_d$, which works adequately well for
low-dimensional problems. When gradients are not available or $g_i$ is
not differentiable, the Bayesian optimisation paradigm exists as an
alternative solution. In this scenario, the training part becomes
slower due to fitting of the surrogate model and the blind
optimisation steps. Nevertheless, the subsequent task of computing the
proposal region $q_i$ becomes faster since $\hat{d}_i$ can be used
instead of $g_i$; hence we avoid to run the simulator
$M_d(\thetab, \vb_i)$ for each query. The
algorithms~\ref{alg:training_GB} and~\ref{alg:training_GP} present the
above procedure.

\subsubsection*{Inference Part}
Performing the inference includes one or more of the following three
tasks; (a) evaluating the unnormalised posterior
$p_{d, \epsilon}(\thetab|\data)$ (b) sampling from the posterior
$ \thetab_i \sim p_{d, \epsilon}(\thetab|\data)$ (c) computing an
expectation $E_{\thetab|\data}[h(\thetab)]$. Computing an expectation
can be done easily after weighted samples are obtained using the
equation~\ref{eq:expectation}, so we will not discuss it separately.

\noindent
Evaluating the unnormalised posterior requires solely the
deterministic functions $g_i$ and the prior distribution $p(\thetab)$;
there is no need for solving the optimisation problems and building
the proposal regions. The evaluation requires iterating over all $g_i$
and evaluating the distance from the observed data. In contrast, using
the GP approach, the optimisation part should be performed first for
fitting the surrogate models $\hat{d}_i(\thetab)$ and evaluate the
indicator function on them. This provides an important speed-up,
especially when running the simulator is computationally
expensive. % The evaluation of the posterior is presented analytically
% in~\ref{alg:posterior_GB} and~\ref{alg:posterior_GP}.

@@

>>>>

Here we present ROMC as a sequence of algorithmic steps. The reader
may think of the algorithmic description as the backbone of the
implementation that will follow. At a high-level, ROMC can be split
into the training and the inference part.

\subsubsection*{Training part}
\noindent
At the training (fitting) part, the goal is the estimation of the
proposal regions $q_i$. The tasks are (a) sampling the nuisance
variables $\vb_i \sim p(\vb)$ (b) defining the optimisation problems
$\min_{\thetab} \: g_i(\thetab)$ (c) obtaining $\thetab_i^*$ (d)
checking whether $d_i^* \leq \epsilon$ and (e) building the bounding
box for obtaining the proposal region $q_i$. If gradients are
available, using a gradient-based method is advised for obtaining
$\thetab_i^*$ much faster. Providing $\nabla_{\thetab} g_i$ in
closed-form provides an upgrade in both accuracy and efficiency; If
closed-form description is not available, approximate gradients with
finite-differences requires two evaluations of $g_i$ for \emph{every}
parameter $\theta_d$, which can only work in low-dimensional
problems. When gradients are not available, the Bayesian optimisation
paradigm exists as an alternative solution. In this scenario, the
training part becomes slower due to fitting of the surrogate model and
the blind optimisation steps. Nevertheless, the subsequent task of
computing the proposal region $q_i$ becomes faster since $\hat{d}_i$
can be used instead of $g_i$. The algorithms~\ref{alg:training_GB}
and~\ref{alg:training_GP} present the above procedure.

\subsubsection*{Inference Part}
Performing the inference includes one or more of the following three
tasks; (a) sampling from the posterior
$ \thetab_i \sim p_{d, \epsilon}(\thetab|\data)$, (b) computing an
expectation $E_{\thetab|\data}[h(\thetab)]$ and (c) evaluating the
unnormalised posterior $p_{d, \epsilon}(\thetab|\data)$. Computing an
expectation can be done easily after weighted samples are obtained
using the equation~\ref{eq:expectation}, so we do not discuss it
separately.

@@

\noindent
Evaluating the unnormalised posterior requires only the
deterministic functions $g_i$ and the prior distribution $p(\thetab)$;
there is no need for solving the optimisation problems and building
the proposal regions. The evaluation requires iterating over all $g_i$
and evaluating the distance from the observed data. In contrast, using
the GP approach, the optimisation part should be performed first for
fitting the surrogate models $\hat{d}_i(\thetab)$ and evaluate the
indicator function on them. This provides an important speed-up,
especially when running the simulator is computationally
expensive.

### 2021-07-19 latex_files/ch3/3_2_2_training.tex
initial draft (without intro and conclusion)

\begin{Code}
---------------------------------- python ----------------------------------
>>> romc.fit_posterior(args*)  # training in a single call
>>> romc.visualize_region(i)   # acceptance region
>>> romc.compute_eps(quantile) # estimates ε
----------------------------------------------------------------------------
\end{Code}

>>>>

\begin{Code}
---------------------------------- python ----------------------------------
>>> romc.fit_posterior(args*)  # training in a single call
>>> romc.visualize_region(i)   # acceptance region
>>> romc.compute_eps(quantile) # estimates eps
----------------------------------------------------------------------------
\end{Code}

### 2021-07-19 latex_files/ch3/3_2_3_inference.tex
initial draft (without intro and conclusion)

  # Summarize the samples (built-in ELFI tool)
  romc.result.summary()
  # Number of samples: 1720
  # Sample means: theta: -0.0792

  # compute expectation
  exp_val = romc.compute_expectation(h=lambda x: np.squeeze(x))
  print("Expected value   : %.3f" % exp_val)
  # Expected value   : -0.079

  exp_var = romc.compute_expectation(h=lambda x: np.squeeze(x)**2)
  print("Expected variance: %.3f" % exp_var)
  # Expected variance: 1.061

>>>>

  # Summarize the samples (built-in ELFI tool)
  romc.result.summary()
  # Number of samples: 19300
  # Sample means: theta: -0.0132

  # compute expectation
  exp_val = romc.compute_expectation(h=lambda x: np.squeeze(x))
  print("Expected value   : %.3f" % exp_val)
  # Expected value   : -0.013

  exp_var = romc.compute_expectation(h=lambda x: np.squeeze(x)**2)
  print("Expected variance: %.3f" % exp_var)
  # Expected variance: 1.116

### 2021-07-19 latex_files/ch4/4_3_exec.tex
initial draft (without intro and conclusion)

In this section, we will present the execution time of the basic ROMC
functionalities. Apart from performing the inference accurately, one
of the notable advantages of ROMC is its efficiency. In terms of
performance, ROMC holds two key advantages.

Firstly, all its subparts are parallelisable; optimising the objective
functions, constructing the bounding boxes, fitting the local
surrogate models, sampling and evaluating the posterior can be
executed in a fully-parallel fashion. Therefore, the speed-up can be
extended as much as the computational resources allow. Specifically,
at the CPU level, the parallelisation can be incorporate all available
cores. A similar design can be utilised in the case of having a
cluster of PCs available. Parallelising the process at the GPU level
is not that trivial, since the parallel tasks are more complicated
than simple floating-point operations. Our current implementation
supports parallelisation only at the CPU level, exploiting all the cores of the CPU.\footnote{At a future update, it is possible to support parallelising the processes using a cluster of PCs.}. In subsection
\ref{subsubsec:parallel}, we will demonstrate the speed-up achieved
through parallelisation.

The second important advantage concerns the execution of the
training and the inference phase in distinct time-slots. Therefore, one
can consume a lot of training time and resources but ask for
accelerated inference. The use of a lightweight surrogate model around
the optimal point exploits this essential characteristic; it trades
some additional computational burden at the training phase, for
exercising faster inference later. In figures \ref{fig:exec_solve},
\ref{fig:exec_regions}, \ref{fig:exec_posterior},
\ref{fig:exec_sample} we can observe this advantage. The
example measured in these figures is the simple 1D example, used in
the previous chapter. We observe that fitting local surrogate models
slows down the training phase (estimating the regions) by a factor of
$1.5$, but provides a speed-up of $15$ at the inference phase i.e.\
approximating unnormalised posterior. This outcome would be even more
potent in larger models, where running the simulator is even more
expensive.

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_solve_grad.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_solve_bo.png}
    \end{center}
    \caption[Execution time for solving the optimisation problems.]{Execution time for defining and solving the optimisation
      problems. We observe that the Bayesian optimisation scheme is much more expensive, performing the task slower by a factor of $75$.}
  \label{fig:exec_solve}
\end{figure}

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_regions_grad.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_regions_grad_fit.png}\\
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_regions_bo.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_regions_bo_fit.png}
    \end{center}
    \caption[Execution time for constructing the n-dimensional bounding box regions.]{Execution time for constructing the n-dimensional
      bounding box region and, optionally, fitting the local surrogate
      models. We observe that fitting the surrogate models incurs a
      small increase by a factor of $1.5$.}
  \label{fig:exec_regions}
\end{figure}

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_posterior_grad.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_posterior_bo.png}\\
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_posterior_grad_fit.png}
    \end{center}
    \caption[Execution time for evaluation the unnormalised posterior.]{Execution time for evaluating the unnormalised posterior
      approximation. We observe that using a Gaussian Process surrogate model is almost 5 times faster than the simulator and the quadratic local surrogate model 15 times faster.}
  \label{fig:exec_posterior}
\end{figure}

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_sample_grad.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_sample_bo.png}\\
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_sample_grad_fit.png}
    \end{center}
    \caption[Execution time for sampling from the approximate posterior.]{Execution time for sampling from the approximate
      posterior. We observe a small speed-up using the local surrogate model.}
  \label{fig:exec_sample}
\end{figure}

\subsubsection{The effect of parallelisation} \label{subsubsec:parallel}

In this section, we measure the execution times for (a) solving the optimisation problems, (b) construct the bounding boxes, (c) sample from the posterior and (d) evaluate the approximate posterior using parallelisation. The model used for this experiment is the simple two-dimensional example of section \ref{subsec:ex1}. All experiments
have been executed in a laptop Dell XPS 15 with an 8-core CPU. Therefore the maximum expected speed-up can reach a factor of 8. Normally the observed speed-up is lower due to the overhead of setting-up the parallel processes. The experiments confirm our general expectation; the parallel version performs all tasks between 2.5 and 6 times faster compared to the sequential, as shown in figures \ref{fig:exec_parallel} and \ref{fig:exec_parallel2}. The most benefited task is solving the optimisation problems which is performed almost 6 times faster. Sampling is executed 3.5 times faster, whereas evaluating the posterior and constructing the bounding boxes almost 2.5 times faster.

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/solve_problems_parallel.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/estimate_regions_parallel.png}
    \end{center}
    \caption[Execution time of the training part exploiting parallelisation]{Execution
      time of the training part exploiting parallelisation. At the left figure, we measure
      the execution time for solving the optimisation problems. At the
      right figure, we measure the execution time for constructing the
      bounding boxes.}
  \label{fig:exec_parallel}
\end{figure}

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/sample_parallel.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/eval_post_parallel.png}
    \end{center}
    \caption[Execution time of the inference part exploiting parallelisation.]{Execution
      time of the inference part exploiting parallelisation. At the left figure, we measure
      the execution time for sampling $n_2=50$ points per region. At the right figure we measure the execution time for evaluating the posterior at a batch of $50$ points.}
  \label{fig:exec_parallel2}
\end{figure}

>>>>

As already stated, it is straightforward for the two basic parts of
the training part to be parallelised. There is a batch of $n_1$
independent objectives functions to be optimised and $n_1$ bounding
boxes to be built. Both tasks can be applied in a parallel utilizing
all the available CPU cores. Our implementation supports
parallelization using the built-in \proglang{Python}
\pkg{multiprocess}. In figure \label{fig:exec_parallel} we observe the
execution times for performing the inference. The parallel version
performs all tasks between 2.5 and 6 times faster compared to the
sequential. Optimisation problems are solved almost 6 times
faster. Sampling is executed 3.5 times faster, whereas evaluating the
posterior and constructing the bounding boxes almost 2.5 times faster.

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/solve_problems_parallel.png}
      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/estimate_regions_parallel.png}\\
      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/sample_parallel.png}
      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/eval_post_parallel.png}

    \end{center}
    \caption[Execution time exploiting parallelisation]{In the first
      line, we compare the parallel and sequential execution of the
      training part and at the second line the inference part. At the
      left figure, we measure the execution time for sampling $n_2=50$
      points per region. At the right figure we measure the execution
      time for evaluating the posterior at a batch of $50$ points.}
  \label{fig:exec_parallel}
\end{figure}

### 2021-07-19 paper.tex
standarized ma2 example
Updates from Overleaf
reordered document
initial draft (without intro and conclusion)

\documentclass[article]{jss}

%% -- LaTeX packages and custom commands ---------------------------------------

@@

%% Vasilis packages
\usepackage{amsmath,amssymb,bbold, xfrac}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{caption}
\usepackage{float}

@@

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo}
\Plaintitle{An Extendable Python Implementation of ROMC}
\Shorttitle{An Extendable \proglang{Python} Implementation of ROMC}

@@

% \begin{leftbar}
% The introduction is in principle ``as usual''. However, it should usually embed
% both the implemented \emph{methods} and the \emph{software} into the respective
% relevant literature. For the latter both competing and complementary software
% should be discussed (within the same software environment and beyond), bringing
p% out relative (dis)advantages. All software mentioned should be properly
% \verb|\cite{}|d. (See also Appendix~\ref{app:bibtex} for more details on
% \textsc{Bib}{\TeX}.)

@@

\begin{leftbar}
For code input and output, the style files provide dedicated environments.
Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
hook. Please make sure that all code is properly spaced, e.g., using
\code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
use ``the usual'' command prompt in the respective software system. For
\proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
the continuation prompt. Generally, comments within the code chunks should be
avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
before and after code input/output should be avoided (see above).
\end{leftbar}

\section{Background}

\subsection{Likelihood-free inference and implicit models}
\input{latex_files/ch2/2_1_LFI_intro.tex}

@@

\section{Example}
\input{latex_files/ch4/4_2_ma2.tex}

@@

The results in this paper were obtained using
\proglang{R}~3.4.1 with the
\pkg{MASS}~7.3.47 package. \proglang{R} itself
and all packages used are available from the Comprehensive
\proglang{R} Archive Network (CRAN) at
\url{https://CRAN.R-project.org/}.

@@

\begin{leftbar}
All acknowledgments (note the AE spelling) should be collected in this
unnumbered section before the references. It may contain the usual information
about funding and feedback from colleagues/reviewers/etc. Furthermore,
information such as relative contributions of the authors may be added here
(if any).
\end{leftbar}

>>>>

\documentclass[article]{jss}
%% -- LaTeX packages and custom commands ---------------------------------------

@@

%% Vasilis packages
\usepackage{amsmath,amssymb,bbold, xfrac}
% \usepackage{xcolor}
\usepackage{multirow}
\usepackage{caption}
\usepackage{float}

@@

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{A Parallelized and Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo}
\Plaintitle{A Parallelized and Extendable Python Implementation of ROMC}
\Shorttitle{A Parallelized and Extendable \proglang{Python} Implementation of ROMC}

@@

% \begin{leftbar}
% The introduction is in principle ``as usual''. However, it should usually embed
% both the implemented \emph{methods} and the \emph{software} into the respective
% relevant literature. For the latter both competing and complementary software
% should be discussed (within the same software environment and beyond), bringing
% out relative (dis)advantages. All software mentioned should be properly
% \verb|\cite{}|d. (See also Appendix~\ref{app:bibtex} for more details on
% \textsc{Bib}{\TeX}.)

@@

% \begin{leftbar}
% For code input and output, the style files provide dedicated environments.
% Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
% or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
% produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
% hook. Please make sure that all code is properly spaced, e.g., using
% \code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
% use ``the usual'' command prompt in the respective software system. For
% \proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
% the continuation prompt. Generally, comments within the code chunks should be
% avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
% before and after code input/output should be avoided (see above).
% \end{leftbar}

\clearpage
\section{Background}
\input{latex_files/ch2/2_1_LFI_intro.tex}

@@

\section{Use-case illustration}
\subsection{Moving average example}
\input{latex_files/ch4/4_2_ma2.tex}

\subsection{Execution time}
\input{latex_files/ch4/4_3_exec.tex}

@@

% The results in this paper were obtained using
% \proglang{R}~3.4.1 with the
% \pkg{MASS}~7.3.47 package. \proglang{R} itself
% and all packages used are available from the Comprehensive
% \proglang{R} Archive Network (CRAN) at
% \url{https://CRAN.R-project.org/}.

@@

% \begin{leftbar}
% All acknowledgments (note the AE spelling) should be collected in this
% unnumbered section before the references. It may contain the usual information
% about funding and feedback from colleagues/reviewers/etc. Furthermore,
% information such as relative contributions of the authors may be added here
% (if any).
% \end{leftbar}
